Disturbance Assimilation

### 2021-06-06 discussion.Rmd
Methods: starting text on RS data & calibration

>>>>

Leftover text from Methods:
  + x SS Multi-Disturbance - distinct
  + x SS Multi-Disturbance w/ distinct, PFT switching
      * mention parallel PFT ensembles
      * need to protect against long-term PFT leakage; related to more general need to "look back"
  + x Simple Multi-site: spatial cov
  + Future
      * SSMD - initially overlapping but diverge
        * how to handle updating of assignment based on new information?
      * SSMD - transition probabilities change with state
        * rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?

### 2021-06-06 methods.Rmd
Methods: starting text on RS data & calibration

  + x SS Multi-Disturbance - distinct
  + x SS Multi-Disturbance w/ distinct, PFT switching
      * mention parallel PFT ensembles
      * need to protect against long-term PFT leakage; related to more general need to "look back"
  + x Simple Multi-site: spatial cov
  + Future
      * SSMD - initially overlapping but diverge
        * how to handle updating of assignment based on new information?
      * SSMD - transition probabilities change with state
        * rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?

## Real-world examples

@@

    + Landtrendr
    http://emapr.ceoas.oregonstate.edu/pages/data/viz/index.html
    + Location:
    tile h03v04
xmin       : -2115585 
xmax       : -1965585 
ymin       : 2564805 
ymax       : 2714805 
Upper Left  (-2115585.000, 2714805.000) (123d 6'32.29"W, 44d41'17.09"N)
Lower Left  (-2115585.000, 2564805.000) (122d35' 7.51"W, 43d23'46.18"N)
Upper Right (-1965585.000, 2714805.000) (121d16'45.91"W, 45d 3'17.45"N)
Lower Right (-1965585.000, 2564805.000) (120d47'15.95"W, 43d45'18.18"N)
Center      (-2040585.000, 2639805.000) (121d56'25.42"W, 44d13'35.07"N)
> res(cover)
[1] 30 30
> crs(cover)
CRS arguments:
 +proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs 
    + Biomass [@Kennedy2018]
      1990-2017
    + Cover
      1990-2017
    + Disturbance
      1984-2011
      - nearest neighbor resampled to same as cover

@@

@@

* Calibrate at US-Me2, 2002-2019 
* Initialize with BADM soil, leaf, and stem C
* Priors (and some fixed parameters) using BADM; default uniform for rest
* Likelihood against tower _annual_ NEE and BADM leaf and stem biomass
* MCMC using DEzs algorithm for 25000 iterations of 2 chains
* GBR, effective sample

>>>>

## Remotely-sensed disturbance in Oregon

Having tested the Multinomial assimilation against simulated data we next applied the algorithm to multiple remotely-sensed disturbance types for an evergreen forested region in the central Oregon Cascades. In summary, we: (1) calibrated VSEM against field data, then (2) used a remotely-sensed disturbance product to construct priors for disturbance rates and PFT transition, then (3) extracted and assimilated ~400 time-series of remotely sensed aboveground biomass, stratified by different disturbance types. We then assessed the ability of the algorithm to detect different disturbances types and magnitudes.

@@

Our study region was a Landsat tile located between 120 47'W and 123 7'W longitude and 43 45'N and 45 '4'N latitude. For this tile we extracted data from three derived Landtrendr datasets (http://emapr.ceoas.oregonstate.edu/pages/data/viz/index.html, tile h03v04):  biomass, cover, and disturbance. All three products were 30m resolution and annual timestep;biomass and cover extended from 1990-2017, while disturbance extended from 1984-2011 and a nearest neighbor algorithm was used to regrid the disturbance data to be the same as biomass and cover. The biomass product ... [@Kennedy2018]. The cover product ... [CITE]. To similify the analysis we focused on a single vegetation type, evergreen forest, which is the dominant vegetation cover in the region and consists primarily of ponderosa pine. The disturbance product ... [CITE]. To simplify analyses we aggregated the disturbance product into five classes: undisturbed, cut, fire, pest, and other.

```{r map, echo=FALSE}
## maps of pre-disturbance biomass, cover, and 2004 disturbance

```

@@

VSEM was calibrated primarily using data from 2002-2019 from the Ameriflux site US-ME2 (lat.lon), a mature Ponderosa pine stand. VSEM pools (leaf, stem, and soil C) were initialized using the earliest available site-specific data from the Ameriflux BADM (Biomass, Ancillary, Disturbance, Management) dataset.

@@

* Calibrate at US-Me2, 2002-2019 
* Priors (and some fixed parameters) using BADM; default uniform for rest
* Likelihood against tower _annual_ NEE and BADM leaf and stem biomass
* MCMC using DEzs algorithm for 25000 iterations of 2 chains
* GBR, effective sample
